---
title: "Coding Challenge 6 - Iterations and Functions"
author: "Derek Anderson"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: gfm
  html_document:
    toc: true
    toc_float: true
  word_document:
  pdf_document:
---
# Global Setup
```{r setup, include=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
library(markdown)
library(drc)
library(dplyr)
# set the default global option for displaying code chunks
knitr::opts_chunk$set(echo = TRUE)
# this sets the root file project to the main R project directory not location of current Rmd file
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Question 1.	

2 pts. Regarding reproducibility, what is the main point of writing your own functions and iterations? 

*Answer: The main point of writing your own functions and iterations is to make analyses repeatable, consistent, and automated. Instead of manually repeating the same code multiple times, a function allows you to define a set of instructions once and apply it to different data inputs. Iterations (such as for loops) allow the same operation to be applied across multiple elements of a dataset. This reduces human error, keeps workflows organized, and ensures that someone else can run the same code and obtain the same results, which is a key principle of reproducible science.*

# Question 2.	

2 pts. In your own words, describe how to write a function and a for loop in R and how they work. Give me specifics like syntax, where to write code, and how the results are returned.

*Answer: A function in R is created using the function() keyword. It allows you to define a reusable block of code that performs a specific task. The basic syntax is:*

*The function name is assigned using <-. Inside the parentheses you define the inputs (arguments). The code inside the curly braces {} contains the operations the function performs. The result is typically returned using return(). Once the function is defined in a script or console, it can be used by calling the function name and supplying input values.*

```{r}
my_function <- function(argument1, argument2) {
  result <- argument1 + argument2
  return(result)
}
#Example:
my_function(2, 3)
```

*A for loop is used to repeat a set of instructions multiple times.*

*The loop variable (i) takes on each value in the sequence (1:5). For every value, the code inside the curly braces {} is executed. For loops are useful when applying the same operation across many items, such as rows in a dataset or elements in a vector.*

*Both functions and loops are written inside R scripts or the console in RStudio, and they help automate repetitive tasks so analyses are easier to reproduce and modify. The syntax is:*

```{r}
for (i in 1:5) {
  print(i)
}
```
# Question 3.

2 pts. Read in the Cities.csv file from Canvas using a relative file path.

```{r}
cities.df <- read.csv("data_files/Cities.csv")
```

# Question 4.	

6 pts. Write a function to calculate the distance between two pairs of coordinates based on the Haversine formula (see below). The input into the function should be lat1, lon1, lat2, and lon2. The function should return the object distance_km. All the code below needs to go into the function.

```{r}
distance_calc <- function(lat1, lon1, lat2, lon2){
# convert to radians
rad.lat1 <- lat1 * pi/180
rad.lon1 <- lon1 * pi/180
rad.lat2 <- lat2 * pi/180
rad.lon2 <- lon2 * pi/180

# Haversine formula
delta_lat <- rad.lat2 - rad.lat1
delta_lon <- rad.lon2 - rad.lon1
a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
c <- 2 * asin(sqrt(a)) 

# Earth's radius in kilometers
earth_radius <- 6378137

# Calculate the distance
distance_km <- (earth_radius * c)/1000

# Return the distance
return(distance_km)
}
```

# Question 5.

5 pts. Using your function, compute the distance between Auburn, AL and New York City 

a.	Subset/filter the Cities.csv data to include only the latitude and longitude values you need and input as input to your function.
b.	The output of your function should be 1367.854 km

```{r}
# filter and subset 
cities.sub <- cities.df %>%
  dplyr::filter(city %in% c("Auburn", "New York")) %>%
  dplyr::select(city, lat, long)
# pass to function
distance_calc(
  lat1 = cities.sub$lat[1],
  lon1 = cities.sub$long[1],
  lat2 = cities.sub$lat[2],
  lon2 = cities.sub$long[2]
)
```

# Question 6.

6 pts. Now, use your function within a for loop to calculate the distance between all other cities in the data. The output of the first 9 iterations is shown below. 

# Bonus point

Have the output of each iteration append a new row to a dataframe, generating a new column of data. In other words, the loop should create a dataframe with three columns called city1, city2, and distance_km, as shown below. The first six rows of the dataframe are shown below.

```{r}
# Get Auburn coordinates
auburn_lat <- cities.df$lat[cities.df$city == "Auburn"]
auburn_long <- cities.df$long[cities.df$city == "Auburn"]

# Create empty dataframe to store results
distance.df <- data.frame(
  City1 = character(),
  City2 = character(),
  Distance_km = numeric(),
  stringsAsFactors = FALSE
)

# Loop through all cities
for (i in 1:nrow(cities.df)) {
  
  dist <- distance_calc(
    lat1 = cities.df$lat[i],
    lon1 = cities.df$long[i],
    lat2 = auburn_lat,
    lon2 = auburn_long
  )
  
  distance.df <- rbind(
    distance.df,
    data.frame(
      City1 = cities.df$city[i],
      City2 = "Auburn",
      Distance_km = dist
    )
  )
}

# View first rows
distance.df
```

# Question 7.

2 pts. Commit and push a gfm .md file to GitHub inside a directory called Coding Challenge 6. Provide me a link to your github written as a clickable link in your .pdf or .docx or .html. 

# Alternative Solutions

Above I used rbind(). You could also loop to create vectors to fill columns: 

## Vector Based

```{r}
# Get Auburn coordinates
auburn_lat <- cities.df$lat[cities.df$city == "Auburn"]
auburn_long <- cities.df$long[cities.df$city == "Auburn"]

# Create your vectors
City1 <- c()
City2 <- c()
Distance_km <- c()

# Perform your loop
for (i in 1:nrow(cities.df)) {

  City1[i] <- cities.df$city[i]
  City2[i] <- "Auburn"

  Distance_km[i] <- distance_calc(
    cities.df$lat[i],
    cities.df$long[i],
    auburn_lat,
    auburn_long
  )
}

# Use vectors to fill data frames 
distance.df2 <- data.frame(City1, City2, Distance_km)

# preview to show that they are the same
head(distance.df2)
```

Lastly. You could use mutate() and select to avoid loops altogether.

## Mutate() Solution

```{r}
# Get Auburn coordinates
auburn_lat <- cities.df$lat[cities.df$city == "Auburn"]
auburn_long <- cities.df$long[cities.df$city == "Auburn"]

# Pipe and Mutate
distance.df3 <- cities.df %>%
  dplyr::mutate(Distance_km = distance_calc(lat, long, auburn_lat, auburn_long)) %>%
  dplyr::select(City1 = city, Distance_km) %>%
  dplyr::mutate(City2 = "Auburn") %>%
  dplyr::select(City1, City2, Distance_km)

# preview to show that they are the same
head(distance.df3)
```

