---
title: "README"
author: "Derek Anderson"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
  md_document:
    variant: gfm
  word_document:
  pdf_document:
---

# README

```{r setup, include=FALSE}
library(knitr)
library(markdown)
library(fs)
```

## Introduction

My name is [Derek Anderson](https://github.com/dbanderson1). I am graduate student in Kinesiology at Auburn University. This repository contains my assignments and practice code for [PLPA 6820 (Reproducible Science)](https://bulletin.auburn.edu/coursesofinstruction/plpa/) taught by [Zach Noel](https://agriculture.auburn.edu/about/directory/faculty/zachary-noel/). Professor Noel is also active on [GitHub](https://github.com/noelzach) and has many great resources for students learning R.

## Overview

Here youвЂ™ll find:

- R scripts for assignments
- Example datasets
- Practice with reproducible workflows (R, Git, GitHub, R Markdown)

## File Links

For ease of reference here are links to graded files:

- assignments (a.k.a. 'coding notes')
  - [assingment_1](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/assignments/assignment_001/assignment_001.R)
  - [assingment_2](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/assignments/assignment_002/assignment_002.R)
  - [assingment_3](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/assignments/assignment_003/assignment_003.R)
  - [assingment_4.md](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/assignments/assignment_004/assignment_004.md), [assingment_4.Rmd](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/assignments/assignment_004/assignment_004.Rmd)
  - [assingment_5.md](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/assignments/assignment_005/assignment_005.md), [assingment_5.Rmd](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/assignments/assignment_005/assignment_005.Rmd)
- coding challenges
  - [coding_challenge_1](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/coding_challenges/coding_challenge_1/coding_challenge_1.R)
  - [coding_challenge_2](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/coding_challenges/coding_challenge_2/coding_challenge_2.R)
  - [coding_challenge_3](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/coding_challenges/coding_challenge_3/coding_challenge_3.R)
  - [coding_challenge_4.md](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/coding_challenges/coding_challenge_4/coding_challenge_4.md), [coding_challenge_4.Rmd](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/coding_challenges/coding_challenge_4/coding_challenge_4.Rmd)

## File Tree

File tree generated using r code and fs():
```{r generate file tree, eval=FALSE}
fs::dir_tree()
```


```{bash file tree, eval=FALSE}
.
в”њв”Ђв”Ђ assignments             # Folder for assignments aka coding notes
в”‚   в”њв”Ђв”Ђ assignment_001
в”‚   в”‚   в””в”Ђв”Ђ assignment_001.R
в”‚   в”њв”Ђв”Ђ assignment_002
в”‚   в”‚   в””в”Ђв”Ђ assignment_002.R
в”‚   в”њв”Ђв”Ђ assignment_003
в”‚   в”‚   в””в”Ђв”Ђ assignment_003.R
в”‚   в”њв”Ђв”Ђ assignment_004
в”‚   в”‚   в”њв”Ђв”Ђ assignment_004.html
в”‚   в”‚   в”њв”Ђв”Ђ assignment_004.md
в”‚   в”‚   в”њв”Ђв”Ђ assignment_004.Rmd
в”‚   в”‚   в”њв”Ђв”Ђ assignment_004_files
в”‚   в”‚   в”‚   в””в”Ђв”Ђ figure-gfm
в”‚   в”‚   в”‚       в””в”Ђв”Ђ include figures-1.png
в”‚   в”‚   в””в”Ђв”Ђ pressure-1.png
в”‚   в””в”Ђв”Ђ assignment_005
в”‚       в”њв”Ђв”Ђ assignment_005.html
в”‚       в”њв”Ђв”Ђ assignment_005.md
в”‚       в”њв”Ђв”Ђ assignment_005.Rmd
в”‚       в””в”Ђв”Ђ assignment_005_files
в”‚           в””в”Ђв”Ђ figure-gfm
в”‚               в”њв”Ђв”Ђ unnamed-chunk-11-1.png
в”‚               в””в”Ђв”Ђ unnamed-chunk-16-1.png
в”њв”Ђв”Ђ coding_challenges       # Folder for Coding Challenges
в”‚   в”њв”Ђв”Ђ coding_challenge_1
в”‚   в”‚   в””в”Ђв”Ђ coding_challenge_1.R
в”‚   в”њв”Ђв”Ђ coding_challenge_2
в”‚   в”‚   в””в”Ђв”Ђ coding_challenge_2.R
в”‚   в”њв”Ђв”Ђ coding_challenge_3
в”‚   в”‚   в””в”Ђв”Ђ coding_challenge_3.R
в”‚   в””в”Ђв”Ђ coding_challenge_4
в”‚       в”њв”Ђв”Ђ coding_challenge_4.html
в”‚       в”њв”Ђв”Ђ coding_challenge_4.md
в”‚       в”њв”Ђв”Ђ coding_challenge_4.Rmd
в”‚       в””в”Ђв”Ђ coding_challenge_4_files
в”‚           в””в”Ђв”Ђ figure-gfm
в”‚               в””в”Ђв”Ђ Commbine and Display Plots-1.png
в”њв”Ђв”Ђ data_files              # Folder with data files used by scripts
в”‚   в”њв”Ђв”Ђ anthro.csv
в”‚   в”њв”Ђв”Ђ BacterialAlpha.csv
в”‚   в”њв”Ђв”Ђ Bull_richness.csv
в”‚   в”њв”Ђв”Ђ corr.csv
в”‚   в”њв”Ђв”Ђ diff_abund.csv
в”‚   в”њв”Ђв”Ђ MycotoxinData.csv
в”‚   в””в”Ђв”Ђ TipsR.csv
в”њв”Ђв”Ђ plpa_6820_assignments.Rproj # Top level R Project
в”њв”Ђв”Ђ README.html
в”њв”Ђв”Ђ README.md               # Top level directory README
в””в”Ђв”Ђ README.Rmd
```

## How to Use

Feel free to browse and copy the code if you are also just learning R and Rmarkdown.

### Clone this repository:
```{bash clone, eval=FALSE}
git clone https://github.com/dbanderson1/plpa_6820_assignments.git
```
