---
title: "README"
author: "Derek Anderson"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: gfm
  html_document:
    toc: true
    toc_float: true
  word_document:
  pdf_document:
---

# README

```{r setup, include=FALSE}
library(knitr)
library(markdown)
library(fs)
```

## Introduction

My name is [Derek Anderson](https://github.com/dbanderson1). I am graduate student in Kinesiology at Auburn University. This repository contains my assignments and practice code for [PLPA 6820 (Reproducible Science)](https://bulletin.auburn.edu/coursesofinstruction/plpa/) taught by [Zach Noel](https://agriculture.auburn.edu/about/directory/faculty/zachary-noel/). Professor Noel is also active on [GitHub](https://github.com/noelzach) and has many great resources for students learning R.

## Overview

Here you’ll find:

- R scripts for assignments
- Example datasets
- Practice with reproducible workflows (R, Git, GitHub, R Markdown)

## File Links

For ease of reference here are links to graded files:

- assignments (a.k.a. 'coding notes')
  - [assingment_1](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/assignments/assignment_001/assignment_001.R)
  - [assingment_2](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/assignments/assignment_001/assignment_002.R)
  - [assingment_3](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/assignments/assignment_001/assignment_003.R)
  - [assingment_4](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/assignments/assignment_001/assignment_004.md)
- coding challenges
  - [coding_challenge_1](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/coding_challenges/coding_challenge_1/coding_challenge_1.R)
  - [coding_challenge_2](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/coding_challenges/coding_challenge_2/coding_challenge_2.R)
  - [coding_challenge_3](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/coding_challenges/coding_challenge_3/coding_challenge_3.R)
  - [coding_challenge_4](https://github.com/dbanderson1/plpa_6820_assignments/blob/main/coding_challenges/coding_challenge_4/coding_challenge_4.md)

## File Tree

File tree generated using r code and fs():
```{r generate file tree, eval=FALSE}
fs::dir_tree()
```


```{bash file tree, eval=FALSE}
.
├── assignments
│   ├── assignment_001
│   │   └── assignment_001.R
│   ├── assignment_002
│   │   └── assignment_002.R
│   ├── assignment_003
│   │   └── assignment_003.R
│   └── assignment_004
│       ├── assignment_004.html
│       ├── assignment_004.md
│       ├── assignment_004.Rmd
│       ├── assignment_004_files
│       │   └── figure-gfm
│       │       └── include figures-1.png
│       └── pressure-1.png
├── coding_challenges
│   ├── coding_challenge_1
│   │   └── coding_challenge_1.R
│   ├── coding_challenge_2
│   │   └── coding_challenge_2.R
│   ├── coding_challenge_3
│   │   └── coding_challenge_3.R
│   └── coding_challenge_4
│       ├── coding_challenge_4.html
│       ├── coding_challenge_4.md
│       ├── coding_challenge_4.Rmd
│       └── coding_challenge_4_files
│           └── figure-gfm
│               └── Commbine and Display Plots-1.png
├── data_files
│   ├── anthro.csv
│   ├── BacterialAlpha.csv
│   ├── Bull_richness.csv
│   ├── corr.csv
│   ├── diff_abund.csv
│   ├── MycotoxinData.csv
│   └── TipsR.csv
├── plpa_6820_assignments.Rproj
├── README.html
├── README.md
└── README.Rmd
```

## How to Use

Feel free to browse and copy the code if you are also just learning R and Rmarkdown.

### Clone this repository:
```{bash clone, eval=FALSE}
git clone https://github.com/dbanderson1/plpa_6820_assignments.git
```
