---
title: "Coding Challenge 6 - Linear Models"
author: "Derek Anderson"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: gfm
  html_document:
    toc: true
    toc_float: true
---

# Question 1.	

4 pts. Read in the data called ŌĆ£PlantEmergence.csvŌĆØ using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor

## Load Libraries
```{r setup, include=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lme4)
library(emmeans)
# install.packages("multcomp")
library(multcomp)
# install.packages("multcompView")
library(multcompView)
# set the default global option for displaying code chunks
knitr::opts_chunk$set(echo = TRUE)
# this sets the root file project to the main R project directory not location of current Rmd file
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
setwd(rprojroot::find_rstudio_root_file())
```

## Read Data
```{r}
plant.emergence <- read.csv("data_files/PlantEmergence.csv")
```

## Create Factors
```{r}
# Factors
plant.emergence$Treatment <- as.factor(plant.emergence$Treatment) #Treatment Factor
plant.emergence$Rep <- as.factor(plant.emergence$Rep) #Rep Factor
plant.emergence$DaysAfterPlanting <- as.factor(plant.emergence$DaysAfterPlanting) #Days After Planting Factor
```

# Question 2.	

5 pts. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. 

```{r}
emergence.lm <- lm(Emergence ~ Treatment * DaysAfterPlanting, data = plant.emergence)
summary(emergence.lm)
anova(emergence.lm)
```

# Question 3.	

5 pts. Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

From the above table we see that "Treatment:DaysAfterPlanting  p = 1>." The interaction is not statistically significant (p=1), meaning the effect of treatment does not depend on the day after planting. Therefore, you can simplify the model and remove the interaction while still keeping the main effects.

```{r}
emergence.lm.simple <- lm(Emergence ~ Treatment + DaysAfterPlanting, data = plant.emergence)
summary(emergence.lm.simple)
anova(emergence.lm.simple)
```

# Question 4.	

5 pts. Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.

```{r}
lsmeans <- emmeans(emergence.lm.simple, ~Treatment) # estimate lsmeans of Treatment
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE) # contrast with Tukey adjustment by default. 
Results_lsmeans
```
# Question 5.	

4 pts. The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 

```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}
plot_cldbars_onefactor(emergence.lm.simple, "Treatment")
```

Treatments that share at least one letter are NOT significantly different from each other.

Treatments that do not share any letters ARE significantly different.

| Treatment | Letter group | Interpretation                                       |
| --------- | ------------ | ---------------------------------------------------- |
| 1         | ab           | not different from treatments labeled **a** or **b** |
| 2         | c            | significantly different from all other treatments    |
| 3         | a            | similar to treatments labeled **a**                  |
| 4         | ab           | overlaps with **a** and **b** groups                 |
| 5         | a            | similar to other **a** treatments                    |
| 6         | a            | similar to other **a** treatments                    |
| 7         | b            | similar to **b** treatments                          |
| 8         | ab           | overlaps **a** and **b**                             |
| 9         | ab           | overlaps **a** and **b**                             |

# Question 6.

2 pts. Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the .md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to your github.

